Probabilistic Network Latency Verification

### 2023-02-13 paper/1_intro.tex
[Paper] Changed abstract, section1-3

\section{Introduction}
The intricate configuraton of network devices is notoriously hard to get right, forcing network 
engineers to accept the reality of Murphy's Law or getting some help from a tool to ensure their 
correctness, namely, network verifier.
Over the last decade, there has been a lot of development in the area of network verification 
generally \cite{hsa}\cite{veriflow}.
While their approach varies a lot (e.g. data plane vs. control plane verification, deterministic vs. 
probabilistic), most of these verifiers shares a common focus, primarily fixated on verifying 
\textit{qualitative} properties -- the convergent functional behavior of a network under various failure 
scenarios -- such as reachability and loop detection.

However, network engineers also have the need to verify \textit{quantitative} properties, such as properties 
related to bandwidth and latency.
Still, barring some recent works about the verification of link load violation \cite{qarc}, work in this area 
remain sparse, despite its undoubtable importance.

In this work, we're exploring the verification of the other side of the network performance metric: latency. 
Certain network deployment often necessitates some latency requirement such as an ISP that has 
latency SLA \cite{Verizon} or deployments of Time-Sensitive Networking (TSN) \cite{TSN}.
We proposed a verification framework to probabilistically verify the latency property of packets 
traversing from a source to destination node under various failure scenarios, by using latency 
measurements of the components in the network.

>>>>

\section{Introduction}
% The intricate configuraton of network devices is notoriously hard to get right, forcing network 
% engineers to accept the reality of Murphy's Law or getting some help from a tool to ensure their 
% correctness, namely, network verifier.
% Over the last decade, there has been a lot of development in the area of network verification 
% generally \cite{hsa}\cite{veriflow}.
% While their approach varies a lot (e.g. data plane vs. control plane verification, deterministic vs. 
% probabilistic), most of these verifiers shares a common focus, primarily fixated on verifying 
% \textit{qualitative} properties -- the convergent functional behavior of a network under various failure 
% scenarios -- such as reachability and loop detection.

Many modern network deployments often necessitate tight performance guarantee in order to sustain 
their increasing demand.
At the same time, end users also have a rising performance standard with regards to their internet browsing 
experience, forcing companies to optimize every stack of their deployment in order to increase conversion.
Hence, network operator must find a way to verify that the network that they configure meets this demand, or 
accept the reality of Murphy's Law.

Over the last decade, there has been a lot of development in the area of network verification 
generally \cite{hsa}\cite{veriflow}.
Still, most of these verifiers are primarily fixated on verifying \textit{qualitative} properties 
-- the convergent functional behavior of a network under various failure scenarios -- 
such as reachability and loop detection, and not \textit{quantitative} properties 
related to performance, such as bandwidth and latency.
Compared to the verification of reachability-based property, verification of bandwidth and latency is 
inherently more challenging since it introduces more dimension to the problem.
In addition to reasoning about the existence of a flow, one needs to reason about the measured 
detail of said flow and model their interaction.

There has been a few notable works that push network verification concept towards this direction \cite{qarc}
\cite{ccac}.
These newer tools model a wide range of additional network behavior that might affect performance to answer a
specific question in mind, namely link-load violation and congestion control behavior.
However, their approaches so far has been focused on worst-case analysis, modeling and answering whether 
an unwanted event could happen or not, which requires further statistical interpretation in order to translate 
them to a percentage guarantee that appear on an SLA.

Moreover, none of these tools were designed to answer questions related to end-to-end latency SLA \cite{Verizon}.
Ideally, such a tool would answer the challenge of modeling various latency altering network behavior, 
such as traffic generation, load balancing scheme, and congestion control, and expresses the important 
statistical property of the resulting end-to-end latency.
Not just in terms of worst-case behaviors and tails, but also the means.

In this work, we propose a verification framework to probabilistically verify the latency property of packets 
traversing from a source to destination node under various failure scenarios, by using latency 
measurements of the components in the network.

### 2023-02-13 paper/3_functional.tex
[Paper] Changed abstract, section1-3

\section{Functional Verification}
For functional property verification, NetDice \cite{netdice} 
has laid the way for verifying reachability between two nodes under failure in a 
probabilistic and efficient manner.
To contextualize our modification, we will briefly describe some parts of 
NetDice's network encoding and exploration algorithm as a prelude to our alteration.
For a more complete complete description of NetDice, please refer to the original 
paper.

@@

In order to preserve some information that will be used in the temporal verification 
stage, we modify this original exploration algorithm to make it more explicit.
We started by formalizing the notion of Equivalence Classes.
An Equivalence Class $\mathcal{E}$ is defined as a 3-tuple $(U_h, D_h, paths)$.
$U_h$ and $D_h$ refers to a set of hot edges that are up and down respectively.
$paths$ refers to the convergent paths between src-dst pair that is produced by the
control plane when the links in $D_h$ fails.
Unlike the original algorithm, we explicitly store $paths$ in $\mathcal{E}$ so that 
it could be used to for the temporal verification stage.

@@

After having the algorithm produced $\mathcal{T}$, we then continue to the temporal verification 
stage.

>>>>

\section{Functional Verification} \label{sec:functional}
For functional verification stage, NetDice \cite{netdice} 
has laid the way for verifying reachability between two nodes under failure in a 
probabilistic and efficient manner.
To contextualize our modification, we will briefly describe some parts of 
NetDice's network encoding and exploration algorithm as a prelude to our alteration.
For a more complete complete description of NetDice, please refer to the original 
paper.

@@

In order to preserve some information that will be used in the temporal verification 
stage, we modify this original exploration algorithm to make it more explicit.
We started by formalizing the notion of Equivalence Classes.
An Equivalence Class $\mathcal{E}$ is defined as a 3-tuple $(U_h, D_h, paths)$.
$U_h$ and $D_h$ refers to a set of hot edges that are up and down respectively.
$paths$ refers to the convergent paths between src-dst pair that is produced by the
control plane when the links in $D_h$ fails.

@@

After each equivalence class produced its convergent paths, we then continue to verify the 
temporal property of that equivalence class in the next stage.

### 2023-02-13 paper/4_temporal.tex
[Paper] Changed abstract, section1-3

\subsection{Latency Label}
Similar to link failure $r$, we use edge-labeling technique to represent latency that 
will get introduced by the connectivity between two nodes. 
Our model will model two significant sources of latency: \textbf{link propagation} 
and \textbf{packet queuing in the node}.
We will encode these latencies by equipping the model with two additional labels.

@@

To encode queuing latency, we first note that the queuing mechanism in modern switches 
usually resides in the output port. %TODO: Cite PISA?
Since a port in a switch is only connected to one other port, we could effectively 
assign the latency to the connectivity between switches.
To do this, we define another function $l_q: E_t \rightarrow \mathcal{D}$.
This distribution signifies the output queue latency in the source node that is 
going to be forwarded to the destination node.
Unlike $l_p$, two symmetrical edges will have two different distributions since they 
represent two different output queues.

In our running example, we will assume that all edges in the topology will have $l_p$
and $l_q$ of gamma distribution with $\lambda = 1$, except $l_q(EZ)$ which will 
have a lognormal distribution with $\alpha = 1$ and $\theta = 1$.

@@

\subsection{Weighted Average and Path Convolution}
Since a convergent behavior of the forwarding plane $paths$ might contain more than 
one paths, we will compute the temporal probability of $paths$ by computing the 
weighted average of the temporal probability of each of the individual path.
The specific weight depends on the load-balancing method used.
%We have implemented the weight computation for ECMP load-balancing scheme.

To get the temporal probability of a single path, we will use the latency information 
from $l_p$ and $l_q$ to compute the latency distribution of the whole path.
To do this, we resort to the methods of \textbf{convolution}.
For each edge $e$ in the path, we will get $l_p(e)$ and $l_q(e)$ and convolve them 
all together to get the path latency distribution.

@@

By introducing the temporal verification algorithm, we're essentially adding additional overhead to the 
exploration algorithm (on top of functional verification) that scales to the size of the Exploration Tree.
In other words, for $\mathcal{T}$ of size $n$, we will need to do temporal verification in each of 
those $n$ Equivalence Class.
We aim to minimize the overhead of temporal verification by introducing some optimization algorithms
later.

\subsection{Numeric Convolution}
However, there is one major problem with a convolution-based technique for computing the latency distribution 
of a given path: not every distribution pair can be convolved analytically.
A closed-form solution of a convolution is usually only available for two distributions of the same type.

>>>>

\subsection{Latency Label}
Similar to link failure $r$, we use edge-labeling technique to represent latency that 
will get introduced by the connectivity between two nodes. 
Our model will model two significant sources of latency: \textbf{link propagation} 
and \textbf{packet queuing} in the node.
We will encode these latencies by equipping the model with two additional labels.

@@

To encode queuing latency, we first note that the queuing mechanism in modern switches 
usually resides in the output port. %TODO: Cite PISA?
Since one port in a switch is only connected to one port in another switch, we could effectively 
assign queuing latency to the connectivity between switches.
To do this, we define another function $l_q: E_t \rightarrow \mathcal{D}$.
This distribution signifies the output queue latency in the source node that is 
going to be forwarded to the destination node.
Unlike $l_p$, two symmetrical edges will have two different distributions since they 
represent two different output queues.

% In our running example, we will assume that all edges in the topology will have $l_p$
% and $l_q$ of gamma distribution with $\lambda = 1$, except $l_q(EZ)$ which will 
% have a lognormal distribution with $\alpha = 1$ and $\theta = 1$

\subsection{Latency Measurement as Distribution}
% To form the required latency distribution

% iid

% independent

@@

\subsection{Weighted Average and Path Convolution}
Since a convergent behavior of the forwarding plane $paths$ might contain more than 
one paths, we will compute the temporal probability of $paths$ by computing the 
weighted average of the temporal probability of each of the individual path.
The specific weight depends on the load-balancing method used.
In our example, we will use ECMP load-balancing scheme.

To get the temporal probability of a single path, we will use the latency information 
from $l_p$ and $l_q$ of each edge in the path to compute the latency distribution of the whole path.
To do this, we resort to the methods of \textbf{convolution}.
For each edge $e$ in the path, we will get $l_p(e)$ and $l_q(e)$ and convolve them 
all together to get the path latency distribution.

@@

At this point, by introducing the temporal verification algorithm, we're essentially adding additional 
overhead to the exploration algorithm (on top of functional verification) that scales to the size of the 
Exploration Tree.
In other words, for $\mathcal{T}$ of size $n$, we will need to do temporal verification in each of 
those $n$ Equivalence Class.
We aim to minimize the overhead of temporal verification by introducing some optimization algorithms
later.

\subsection{Numeric Convolution}
There is one major problem with a convolution-based technique for computing the latency distribution 
of a given path: not every distribution pair can be convolved analytically.
A closed-form solution of a convolution is usually only available for two distributions of the same type.

### 2023-02-13 paper/main.tex
[Paper] Changed abstract, section1-3

\begin{document}
\begin{abstract}
    To combat configuration errors that a given network might have, network verifiers have emerged as 
    one of the promising solution. 
    State-of-the-art network verifiers mainly focused on evaluating qualitative properties under 
    various scenarios of network failures, such as reachability and loop detection.

    However, as modern networks evolved and performance need becomes more stringent -- often 
    expressed in terms of Service Level Agreement (SLA) -- there is a need to evolve network 
    verifiers to also reason about quantitative performance properties. 
    Works in quantitative network verifiers that has arose in recent years mainly focused on one 
    side of the network performance metric: bandwidth and load violation properties. 
    Questions about the other side of network performance metric, latency, were left unanswered.

    In this work, we introduce a verification framework, \tool, that can answer the probability of 
    a given temporal property being true given latency distributions of individual links and 
    nodes in the network. 
    We use the convolution operation on these probability distributions to calculate the end-to-end 
    path latency distribution, and answer property by analyzing the statistical property of this 
    resulting distribution.
    We also introduce two effective optimization techniques to exploit the symmetry in the verification 
    process to reduce the amount of convolution operation needed.

    Our evaluation shows that \tool can verify temporal property 4 orders of magnitude faster on average than 
    our baseline, which is devoid of all optimization.
    We also identify that highly symmetrical topology, like Fat Tree, benefits from our optimization the most 
    and can cut the amount of convolution up to 6 orders of magnitude.
\end{abstract}

>>>>

\begin{document}
\begin{abstract}
    As modern networks evolved and performance need became more stringent, often 
    expressed in terms of Service Level Agreement (SLA), there is a rising need to verify that these 
    agreements hold true.
    Designing a network verifier that can practically reason about quantitative properties proves to be 
    a unique challenge, since it needs to be able to expressively model many quantitative phenomenon, like 
    congestion control and load balancing scheme, while being reasonably performant and not 
    requiring too many inputs from its user. 
    In this work, we introduce a network latency verification framework, \tool, that could probablistically 
    verify temporal properties of a network given latency measurements of its components. 
    We also introduce two effective optimization techniques to exploit the symmetry in the verification 
    process to significantly reduce the verification runtime.
    Our evaluation shows that \tool\ can verify temporal property 90\% faster on average than 
    our baseline, which is devoid of all optimization.
    We also identify that highly symmetrical topology, like Fat Tree, benefits from our optimization the most 
    and can cut the amount of convolution up to 6 orders of magnitude.
\end{abstract}
